Bayesian exponential random graph modelling of KWL Burial network

### 2020-04-01 Note.Rmd
paste some paragraphs from dissertation proposal

>>>>

# Testing the predictions of the models

My hypothesis is that if European colonization created monopolies for trade between a small number of indigenous people, and this resulted in the shift from a more corporate to a more network organization, then I would expect to see changes in mortuary practices, food consumption and feasting behaviors, and social network organization at Kiwulan.

Before European contact, households consisted of single and multi-family units and were organized in a corporate mode with limited economic inequality primarily based on age, gender, or different corporate groups. I expect that burial goods will reveal a relatively even distribution of the quantity and quality of items. Social network analysis will indicate multiple small groups of similar size. If there is unequal distribution of prestige or high quality goods, then it will be restricted to group who share similar demographic characteristics such as age. Production and consumption of ceramics will be based on corporate groups. Thus, diversity of sources and style of pottery within a family unit will be more similar than between family units. Activity areas of family units will reflect more even distribution of high ranked subsistence resources, and cache pits will present less variability in size and contents.

After European contact, there will be more obvious inequalities in the distribution of prestige materials in burials and settlement areas, indicative of an increase in the concentration of wealth at Kiwulan. Prestige goods will not only appear in adult burials but also child burials, which reflect wealth accumulation within a family. Higher quality imported goods will mainly appear in a few but not all individual burials because of control of long distance trade. Social network analysis of burials will present a higher density and higher degree of centrality based on individual power. Production and consumption of ceramics will extend beyond the corporate groups and present more standardized sizes and shapes, which show an emphasis on production for feasting, and control of production by a small number of individuals. Activity areas will reflect uneven distribution of high ranked subsistence resources, and cache pits will present higher variability in size and contents.

### 2020-04-01 analysis/paper/paper.Rmd
paste some paragraphs from dissertation proposal

The burials we analyzed in this study are from Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan that experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century. The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data and diagnostic porcelain as time indicators.

>>>>

The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan that experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century. The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans. The burials were located in the north margin of the AD section which is the largest open area at Kiwulan site.

The archaeological evidence in the Yilan plain indicates some elements of social complexity around 17th century, including increasingly sedentary settlements, cultivation of rice, complex inter- and intra-trade networks, the use of prestige goods, and differential burial practices (Chen 2004; Cheng 2008; Li 2014). Previous studies of social organization of the Kiwulan site suggest two different interpretations of this evidence. Cheng (2008) interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy. However, Hsieh (2012) suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods. She finds that the rich burials were usually associated with elders, which might indicate achieved, rather than inherited, status. The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese (Borao Mateo 2009; Li and Wu 2006). If Cheng is correct and socio-economic inequality was present, it might be rooted in the use of prestige goods traded in the area before the Europeans arrived. In this case, pre-existing inequalities might have become more accentuated as a response to European contact. Did the degree of economic inequality change over time, and how did European contact impact indigenous social organization? These questions motivate this project.

The archaeological materials include around 139,900 pottery sherds, 12,500 imported ceramics and stonewares, 1,100 wooden artifacts, 730 stone tools, 425 metal artifacts, 7,400 glass beads and agates beads, 100 pipes, several ornaments made of animal bones, shells, and wood, and a faunal assemblage (Chen 2007). In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated (Figure 5). These postholes were found aligned in a north-south direction with construction marks, which were interpreted as the remains of house structures. Moreover, the area with burials is oriented in an east-west direction located on the north side of the house structures (Chen 2007). Previous research indicates the presence of economic inequality between burials, but there is debate about whether these represent an egalitarian or stratified society (Cheng 2008; Hsieh 2009a). In the following section I explain the model I will use to evaluate these claims for social inequality, and how I will determine the likely causes of any changes over time that I identify.

The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials (Chen 2005; Cheng 2008), which might indicate the pursuit of prestige or wealth. In addition to prestige, European colonial powers could have provided military support to indigenous allies. For example, Voss’s (2005) study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local indigenous people. A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals. In this study, I will use this model of competition to explain the changes in social inequality at Kiwulan before and after European contact.
